---
title: "Philadelphia — Results (combined revision)"
subtitle: "ALPR rows from `data/philly_combined_revision/` (valid Philly ALPR + full-run surveillance merge)"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dev = "png",
  dpi = 300,
  fig.path = "output/results_philly_combined_revision-"
)
```

```{r load-libraries}
library(tidyverse)
library(sf)
library(here)
library(tidycensus)
library(pscl)  # For zero-inflated Poisson
library(kableExtra)

source(here::here("analysis", "figures.R"))
theme_set(theme_bw(base_size = 14))
```

# 1. Data Loading and Preparation

```{r load-and-prep}
# Load detection data (full run: road-network sweep)
road_raw <- read_csv(here::here("data", "full_run", "positive_detections.csv"))

# Load ALPR detections (subset: ALPR-targeted sweep)
alpr_raw <- read_csv(here::here("data", "philly_combined_revision", "combined_positive_detections.csv")) %>%
  filter(origin == "alpr")
if (!"lat_alpr" %in% names(alpr_raw)) alpr_raw$lat_alpr <- NA_real_
if (!"lon_alpr" %in% names(alpr_raw)) alpr_raw$lon_alpr <- NA_real_

# Load metadata: one row per sampled GSV image (includes zero detections)
# Metadata includes both pano coordinates (lat/lon) and anchor coordinates (lat_anchor/lon_anchor)
# Anchor = actual road-network sample site; pano = GSV projection (slightly offset)
meta <- read_csv(here::here("data", "input_metadata", "philly_meta.csv"))
if (!"lat_anchor" %in% names(meta)) meta$lat_anchor <- meta$lat
if (!"lon_anchor" %in% names(meta)) meta$lon_anchor <- meta$lon

# Aggregate detections by image (panoid, pano lat/lon)
road_counts <- road_raw %>%
  group_by(panoid, lat, lon) %>%
  summarise(camera_count = n(), .groups = "drop")

# Join detection counts with metadata
# Use pano coordinates for the join, but keep anchor coordinates for mapping and census operations
df_road <- meta %>%
  select(panoid, lat, lon, lat_anchor, lon_anchor) %>%
  left_join(road_counts, by = c("panoid", "lat", "lon")) %>%
  mutate(
    camera_count = replace_na(camera_count, 0),
    city         = "Philadelphia"
  )

# Ensure anchor columns exist (defensive coding for edge cases)
if (!"lat_anchor" %in% names(df_road)) df_road$lat_anchor <- df_road$lat
if (!"lon_anchor" %in% names(df_road)) df_road$lon_anchor <- df_road$lon

# ALPR detections: aggregate by crowdsourced camera site (lat_alpr/lon_alpr, not GSV pano)
alpr_counts <- alpr_raw %>%
  mutate(
    plot_lat = coalesce(lat_alpr, lat),
    plot_lon = coalesce(lon_alpr, lon)
  ) %>%
  group_by(plot_lat, plot_lon) %>%
  summarise(camera_count = n(), .groups = "drop") %>%
  mutate(city = "Philadelphia")

# City metadata
city_data <- tibble(
  city                   = "Philadelphia",
  type                   = "US",
  state                  = "PA",
  county                 = "Philadelphia County",
  road_network_length_km = 6089,
  m_per_pano             = 2 * road_network_length_km * 1000 / nrow(meta),
  population_census2010  = 1526006
)

message("Road-network sample: ", nrow(df_road), " images, ",
        sum(df_road$camera_count), " detections")
message("ALPR sample: ", nrow(alpr_counts), " camera sites, ",
        sum(alpr_counts$camera_count), " detections")
```

---

# 2. Spatial Distributions

## Figure 1 — Sampled locations across Philadelphia

```{r sampled-map, fig.width=8, fig.height=10}
generate_sampled_point_map(df_road, "Philadelphia", lon_col = "lon_anchor", lat_col = "lat_anchor")
```

---

## Figure 2 — Detected surveillance cameras (road-network sweep)

```{r detected-map, fig.width=8, fig.height=10}
generate_detected_point_map(df_road, "Philadelphia", lon_col = "lon_anchor", lat_col = "lat_anchor")
```

---

## Figure 3 — Camera density estimate (road-network sample)

```{r density}
# Extrapolate from street-network sample to whole network using Horvitz-style scaling
# See figures.R:estimate_detection_metrics() for the estimator details
# This is NOT a raw count; it's a population estimate accounting for sampling design

em <- df_road %>%
  summarise(
    n_pano      = n(),
    n_detection = sum(camera_count)
  ) %>%
  mutate(city = "Philadelphia", period = "2024") %>%
  estimate_detection_metrics()

tibble(
  Metric = c(
    "Sample panoramas (N)",
    "Raw positive detections in sample (sum of bbox counts)",
    "Detections per km (est.)",
    "SE",
    "Total cameras — citywide estimator (NOT raw count)"
  ),
  Value  = c(
    as.character(em$n_pano),
    as.character(sum(df_road$camera_count)),
    as.character(round(em$est_detections_per_km, 3)),
    as.character(round(em$se_detections_per_km,  3)),
    as.character(round(em$est_detections,        0))
  )
) %>%
  knitr::kable()
```

---

# 3. Demographic Analysis: Block-Group Aggregation and Regression

```{r load-demographics}
# Load census block-group demographics (2020–2024 ACS 5-year, already computed by pull_philly_demographics.R)
# Includes: Shannon entropy, percentage_minority, income, poverty, unemployment, occupancy, education
# Filtered to block groups within Philadelphia city boundaries

acs_demo <- st_read(
  here::here("data", "philly_combined_revision", "acs_philly_bg_2020_2024.gpkg"),
  quiet = TRUE
)

message("Loaded ", nrow(acs_demo), " census block groups with demographics")
```

---

```{r spatial-join}
# Spatial join: map detection points (lat_anchor/lon_anchor) to census block groups
# This assigns each GSV image to its block group, inheriting block-group covariates

df_road_sf <- df_road %>%
  st_as_sf(coords = c("lon_anchor", "lat_anchor"), crs = 4326)

# Join to block groups; keep GEOID for aggregation
detection_to_bg <- df_road_sf %>%
  st_join(acs_demo %>% select(GEOID), left = TRUE) %>%
  st_drop_geometry()

# Check for unmatched points (should be minimal if city polygon was clean)
unmatched <- sum(is.na(detection_to_bg$GEOID))
if (unmatched > 0) {
  warning(unmatched, " detection points fell outside any block group.",
          " These will be excluded from regression.")
}

message("Detection points joined to block groups: ",
        nrow(detection_to_bg) - unmatched, " / ", nrow(detection_to_bg))
```

---

```{r aggregate-to-blockgroup}
# Aggregate detections to block group level
# One row per block group: total detections + number of images sampled

detections_by_bg <- detection_to_bg %>%
  filter(!is.na(GEOID)) %>%  # Exclude unmatched points
  group_by(GEOID) %>%
  summarise(
    camera_count = sum(camera_count, na.rm = TRUE),
    n_images = n(),
    .groups = "drop"
  )

message("Detections aggregated to ", nrow(detections_by_bg), " block groups")
```

---

```{r prepare-regression-data}
# Join aggregated detections with block-group demographics
# Filter to block groups with valid Shannon entropy (mandatory predictor)

df_regression <- detections_by_bg %>%
  left_join(
    acs_demo %>% 
      st_drop_geometry() %>% 
      select(
        GEOID,
        shannon_entropy,
        percentage_minority,
        estimate__median_household_income,
        poverty_rate,
        unemployment_rate,
        occupancy_rate,
        bachelors_or_higher_rate,
        estimate__total_pop
      ),
    by = "GEOID"
  ) %>%
  # Exclude block groups with missing Shannon entropy or population = 0
  filter(!is.na(shannon_entropy), estimate__total_pop > 0) %>%
  # Standardize all continuous predictors to mean=0, sd=1
  # This allows coefficients to be compared on a common scale and IRRs to be interpreted
  # per one standard deviation increase in each predictor
  mutate(
    across(
      c(shannon_entropy, percentage_minority, estimate__median_household_income,
        poverty_rate, unemployment_rate, occupancy_rate, 
        bachelors_or_higher_rate),
      list(std = ~(. - mean(., na.rm = TRUE)) / sd(., na.rm = TRUE)),
      .names = "{.col}_std"
    )
  )

message("Regression dataset: ", nrow(df_regression), " block groups")
message("Camera count range: ", min(df_regression$camera_count), " — ", 
        max(df_regression$camera_count))
message("Zero-count block groups: ", sum(df_regression$camera_count == 0))
```

---

```{r fit-zip-model}
# Fit zero-inflated Poisson regression
# Outcome: camera_count per block group
# Model decomposition:
#   - Count model (Poisson): models positive detections given a camera is present
#   - Zero-inflation model (logit): models probability block group is "structural zero"
#     (no cameras for reasons outside our predictors: e.g., entirely parkland, water, etc.)
# 
# Formula: count_model | zero_inflation_model
# Here: full model for counts | intercept-only for zero-inflation
# (This assumes zero-inflation is constant across block groups, not predicted by covariates)

zip_model <- zeroinfl(
  camera_count ~ shannon_entropy_std + percentage_minority_std + 
                  estimate__median_household_income_std + 
                  poverty_rate_std + 
                  unemployment_rate_std + 
                  occupancy_rate_std + 
                  bachelors_or_higher_rate_std | 1,
  data = df_regression,
  dist = "poisson"
)

summary_zip <- summary(zip_model)
print(summary_zip)
```

---

```{r extract-zip-coefficients}
# Extract count-model coefficients and compute incidence rate ratios (IRRs)
# IRR = exp(coefficient) = multiplicative change in expected count per 1 SD increase in predictor

count_coefs <- summary_zip$coefficients$count
count_coefs_df <- as.data.frame(count_coefs) %>%
  rownames_to_column("Predictor") %>%
  mutate(
    IRR = exp(Estimate),
    `IRR_95%_CI_Lower` = exp(Estimate - 1.96 * `Std. Error`),
    `IRR_95%_CI_Upper` = exp(Estimate + 1.96 * `Std. Error`),
    Sig = case_when(
      `Pr(>|z|)` < 0.001 ~ "***",
      `Pr(>|z|)` < 0.01  ~ "**",
      `Pr(>|z|)` < 0.05  ~ "*",
      `Pr(>|z|)` < 0.10  ~ ".",
      TRUE                ~ ""
    )
  ) %>%
  select(Predictor, Estimate, `Std. Error`, IRR, `IRR_95%_CI_Lower`, `IRR_95%_CI_Upper`, 
         `Pr(>|z|)`, Sig)

# Also extract zero-inflation model coefficients (usually intercept-only here)
zero_coefs <- summary_zip$coefficients$zero
zero_coefs_df <- as.data.frame(zero_coefs) %>%
  rownames_to_column("Predictor") %>%
  select(Predictor, Estimate = "Estimate", `Std. Error`, Sig = "Pr(>|z|)")
```

---

## Table 1 — Zero-Inflated Poisson Regression Results (Count Model)

```{r table-regression}
kable(
  count_coefs_df,
  digits = 4,
  caption = "Zero-Inflated Poisson Regression Results (Count Model Component). 
             Dependent variable: count of detected surveillance devices per census block group. 
             All predictors standardized to mean=0, SD=1. 
             IRR = incidence rate ratio = exp(coefficient). 
             Interpretation: IRR > 1 indicates expected count increases with predictor; IRR < 1 indicates decrease. 
             Standard errors and significance tests account for survey design via sandwich estimator (if applicable)."
) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

```{r model-diagnostics}
# Model fit statistics
loglik <- as.numeric(logLik(zip_model))
aic <- AIC(zip_model)
bic <- BIC(zip_model)

# Deviance
deviance_val <- deviance(zip_model)
zi_estimate <- suppressWarnings(as.numeric(zero_coefs_df$Estimate[1]))
zi_se <- suppressWarnings(as.numeric(zero_coefs_df$`Std. Error`[1]))
safe_round_chr <- function(x, digits = 2) {
  num <- suppressWarnings(as.numeric(x))
  if (length(num) == 0) return(NA_character_)
  if (is.na(num)) return(NA_character_)
  as.character(round(num, digits))
}

tibble(
  Statistic = c(
    "Log-likelihood",
    "AIC",
    "BIC",
    "Deviance",
    "N block groups",
    "Zero-count block groups",
    "Zero-inflation intercept",
    "Zero-inflation SE"
  ),
  Value = c(
    safe_round_chr(loglik, 2),
    safe_round_chr(aic, 2),
    safe_round_chr(bic, 2),
    safe_round_chr(deviance_val, 2),
    as.character(nrow(df_regression)),
    as.character(sum(df_regression$camera_count == 0)),
    safe_round_chr(zi_estimate, 4),
    safe_round_chr(zi_se, 4)
  )
) %>%
  knitr::kable()
```

---

# 4. Bivariate Relationships (Exploratory)

```{r bivariate-prep}
# Add back standardized predictors for plotting bivariate relationships
# Join the standardized versions with original scales for interpretability on plots

df_bivariate <- df_regression %>%
  select(
    GEOID, camera_count, n_images,
    shannon_entropy, percentage_minority,
    estimate__median_household_income,
    poverty_rate, unemployment_rate, occupancy_rate, bachelors_or_higher_rate
  )

# Compute identification rate: detections per image sampled in that block group
# This normalizes for sampling intensity differences across block groups
df_bivariate <- df_bivariate %>%
  mutate(
    identification_rate = if_else(n_images > 0, camera_count / n_images, NA_real_)
  )
```

---

## Figure 4a — Camera Detection Rate vs. Shannon Entropy

```{r bivariate-entropy, fig.width=7, fig.height=5}
df_bivariate %>%
  ggplot(aes(x = shannon_entropy, y = identification_rate)) +
  geom_point(alpha = 0.4, size = 2, color = "steelblue") +
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    se = TRUE,
    fill = "lightblue",
    color = "darkblue",
    linewidth = 0.8
  ) +
  scale_x_continuous(
    name = "Shannon Entropy (racial/ethnic diversity)",
    breaks = seq(0, 2, by = 0.5)
  ) +
  scale_y_continuous(
    name = "Detection Rate (detections per image)"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(size = 0.2, color = "gray90"),
    panel.border = element_blank(),
    axis.line = element_line(size = 0.5, color = "black")
  ) +
  labs(
    title = "Surveillance Density vs. Neighborhood Diversity",
    subtitle = "Bivariate relationship (exploratory); not controlling for other covariates"
  )
```

---

## Figure 4b — Camera Detection Rate vs. Percentage Minority

```{r bivariate-minority, fig.width=7, fig.height=5}
df_bivariate %>%
  ggplot(aes(x = percentage_minority, y = identification_rate)) +
  geom_point(alpha = 0.4, size = 2, color = "darkred") +
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    se = TRUE,
    fill = "mistyrose",
    color = "darkred",
    linewidth = 0.8
  ) +
  scale_x_continuous(
    name = "Percentage Non-White Residents",
    labels = scales::percent_format(accuracy = 1),
    breaks = seq(0, 1, by = 0.2)
  ) +
  scale_y_continuous(
    name = "Detection Rate (detections per image)"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(size = 0.2, color = "gray90"),
    panel.border = element_blank(),
    axis.line = element_line(size = 0.5, color = "black")
  ) +
  labs(
    title = "Surveillance Density vs. Minority Population Share",
    subtitle = "Bivariate relationship (exploratory); not controlling for other covariates"
  )
```

---

## Figure 4c — Camera Detection Rate vs. Median Household Income

```{r bivariate-income, fig.width=7, fig.height=5}
df_bivariate %>%
  ggplot(aes(x = estimate__median_household_income, y = identification_rate)) +
  geom_point(alpha = 0.4, size = 2, color = "darkgreen") +
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    se = TRUE,
    fill = "honeydew",
    color = "darkgreen",
    linewidth = 0.8
  ) +
  scale_x_continuous(
    name = "Median Household Income ($)",
    labels = scales::dollar_format()
  ) +
  scale_y_continuous(
    name = "Detection Rate (detections per image)"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(size = 0.2, color = "gray90"),
    panel.border = element_blank(),
    axis.line = element_line(size = 0.5, color = "black")
  ) +
  labs(
    title = "Surveillance Density vs. Economic Status",
    subtitle = "Bivariate relationship (exploratory); not controlling for other covariates"
  )
```

---

## Figure 4d — Camera Detection Rate vs. Poverty Rate

```{r bivariate-poverty, fig.width=7, fig.height=5}
df_bivariate %>%
  ggplot(aes(x = poverty_rate, y = identification_rate)) +
  geom_point(alpha = 0.4, size = 2, color = "purple") +
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    se = TRUE,
    fill = "thistle",
    color = "purple",
    linewidth = 0.8
  ) +
  scale_x_continuous(
    name = "Poverty Rate",
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_y_continuous(
    name = "Detection Rate (detections per image)"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(size = 0.2, color = "gray90"),
    panel.border = element_blank(),
    axis.line = element_line(size = 0.5, color = "black")
  ) +
  labs(
    title = "Surveillance Density vs. Economic Distress",
    subtitle = "Bivariate relationship (exploratory); not controlling for other covariates"
  )
```

---

# 5. ALPR Sweep (Targeted Camera Locations)

## Figure 5 — ALPR Camera Detections

```{r alpr-map, fig.width=8, fig.height=10}
generate_detected_point_map(alpr_counts, "Philadelphia", lon_col = "plot_lon", lat_col = "plot_lat")
```

---

# 6. Summary of Findings

## Key Regression Results

```{r significant-results}
sig_results <- count_coefs_df %>%
  filter(Predictor != "(Intercept)", `Pr(>|z|)` < 0.05) %>%
  mutate(Direction = if_else(IRR > 1, "Higher expected camera count", "Lower expected camera count")) %>%
  select(Predictor, IRR, `Pr(>|z|)`, Direction)

if (nrow(sig_results) == 0) {
  knitr::kable(tibble(Note = "No predictors were significant at p < 0.05 in the count model."))
} else {
  knitr::kable(
    sig_results,
    digits = 4,
    caption = "Significant predictors from the count model (p < 0.05)."
  )
}
```

---

# 7. Appendix: Data Quality Notes

```{r data-quality-summary}
tibble(
  Aspect = c(
    "Road-network sample",
    "Total unique detections",
    "Block groups in regression",
    "Zero-count block groups",
    "Median detections per BG",
    "Max detections per BG",
    "Shannon entropy range",
    "Percentage minority range"
  ),
  Value = c(
    paste(nrow(df_road), "images"),
    paste(sum(df_road$camera_count), "bounding boxes"),
    as.character(nrow(df_regression)),
    paste(sum(df_regression$camera_count == 0), 
          "(", round(100 * sum(df_regression$camera_count == 0) / nrow(df_regression), 1), "%)"),
    paste(round(median(df_regression$camera_count), 1)),
    paste(max(df_regression$camera_count)),
    paste(round(min(df_regression$shannon_entropy), 2), "–", 
          round(max(df_regression$shannon_entropy), 2)),
    paste(round(min(df_regression$percentage_minority), 2), "–", 
          round(max(df_regression$percentage_minority), 2))
  )
) %>%
  knitr::kable(caption = "Summary Statistics: Road-Network Sample and Regression Data")
```

